\vspace{-4em}Rethinking Model Architecture and Means of Delivery to get Video Interaction with Generative Neural Networks Online

### 2020-11-22 writing/paper/draft_02/draft_02.tex
Gregs tweaks

\begin{abstract}
	% 250 Words
	The lines between scientific researcher and artist are continually being blurred as artists adopt machine learning methods\cite{anna_ridler}, or researchers apply their tools in novel and interesting ways\cite{deep_dream_guy} rethinking how cutting-edge research can be applied to create image generation tools or interrogate the tools themselves. \textit{Learning to See} by Memo Akten\cite{lts_memo} is one such seminal artwork which does both. Reframing the tools which we are in fact living with day-to-day in the context of the gallery provides an opportunity to think deeper about how a neural network interprets the data it is given and witness the capabilities and limitations of such a system. \textit{Learning to See} in particular has a uniquely simple interaction allowing the visitor to play with the network in real time. This work shows that simply shrinking the Pix2Pix\cite{1611.07004} model which powers \textit{Learning to See}, and a custom made WebGL wrapper for TensorflowJS to make a pipeline more suited to video achieves useable results when served online, even on commonplace integrated graphics processing units. This brings the unique interaction developed by Akten \textit{et al}\cite{2003.00902} into the homes of many people and offers exciting possibilites to future video based interaction with machine learing on the internet. The work is currently online here: \url{https://learning-to-learn-to-see.netlify.app/}, and code available here: \url{https://github.com/joshmurr/cci-pix2pix-demo/}.
\end{abstract}

@@

\section{Introduction}
% 500 Words
% --- 568 WORDS ---
Machine Learning (ML) is a part of our everyday lives. For the technical among us it is currently easy enough to see the patterns in the outputs of these black-box systems; as the end of our sentence is predicted as we type, or an item we did not know we wanted is offered to us as we shop. There are however many non-technical users of ML systems who can only see the black-box, or perhaps nothing at all; as Arthur C. Clarke famously said \textit{``Any sufficiently advanced technology is indistinguishable from magic.''}\cite{toffler_1972} It is often the case that exciting new technologies are adopted by artists as readily as industry because the gallery provides a unique environment to present back the to the user one of these black-boxes in a drastically different context. Artists such as Mario Klingemann\cite{mario_klingemann}, Terence Broad\cite{terry_broad} or Anna Ridler\cite{anna_ridler} are all excellent examples of artists who have done just this; delivering the \textit{medium as the message}.

\textit{Learning to See} by Memo Akten\cite{lts_memo} is another example, but unique in its own way in that it gives the user an interactive experience of the artwork; an immediate image translation of arbitrary objects on a tabletop to a rich image of waves crashing, the night sky or ever changing flowers. The user is able to manipulate the input data as they move the objects on the table and witness how the model interprets this data to produce an output image projected in front of them. Although the model still exists as a black box, being able to see how input-affects-output in real time is extremely powerful.

@@

Unfortunately the gallery in itself is a black box to many, entry restricted by a fee or simply by the location of the gallery. Moving such interactive experiences online is a clear next step as once something is online it is immediately available to anyone with an internet connection --- this then poses numerous more challenges. Even a modestly sized NN can contain many millions of parameters. Modern computers are able to perform millions of floating point operations per second (FLOPS) but advances in ML research have shown that \textit{deeper} models and more data produces better results\cite{2005.14165}; as such, without expensive hardware acceleration from discrete graphics processing units (GPU) or even more modern tensor processing units (TPU), an average modern computer cannot crunch the numbers needed to run a large, deep NN\@. Even if one \textit{could} perform the necessary FLOPS, storing this number of parameters can take up gigabytes of memory and so downloading the weights and biases of a pre-trained model can be a time consuming process detracting from the real-time focus.

@@

\textit{Learning to See} is an adaptation of the Pix2Pix model\cite{1611.07004} which is a Conditional Generative Adversarial Network (cGAN)\cite{1701.00160}. Unlike a conventional GAN\cite{1406.2661}, a cGAN takes in two data samples as input, in the case of the Pix2Pix model it is a \textit{target} image and an \textit{input} image. Pix2Pix also learns a structured loss which means it is able to penalize any possible structure that differs between output and target\cite{1611.07004}. This model architecture is extremely flexible and has been shown to be useful for many use cases as detailed in the original paper. However neither the original Pix2Pix paper or \textit{Learning to See} had an incentive to create a small model; a good result is simply a sharp image with a predictable output from a given input. As stated by the Pix2Pix authors measuring the results of an image generating GAN is an open issue\cite{1606.03498, 1611.07004} and in an artistic context quantifiying the outcome is largely pointless. A good result is one which delivers the message of the artist, thus if a meaningful interaction can be achieved then perhaps some sacrifices can be made to improve the speed and reduce the size of the model.

@@

Designing and training a NN is of course not limited to the big name ML frameworks like Tensorflow\cite{tf} or PyTorch\cite{pytorch}, but given the modular nature of a NN and the want in most cases to utilise the GPU for faster training and inference, it is no surprise that a limited number of ML frameworks have become the go-to to create and iterate on a novel ML model and to take care of interfacing with a highly optimized backend and the GPU at a lower level. Tensorflow offers a relatively full end-to-end pipeline from designing and training a model in Python, to saving and converting a model into something can be run in the browser using TensorflowJS\cite{tfjs}. A trained model is simply a collection of weights and biases and to use this model on a different device or context, these weights and biases need to be loaded into a model of the same architecture as was used to train the model. The pipeline from training-to-browser is largely about keeping data structures in order and consistent such that the next process knows how to work with it. As part of the process Tensorflow as developed the TensorflowJS Converter\cite{tfjs_converter} which creates a \verb|model.json| file which is the blueprints of the model, and a series of binary files which hold the weights. The pipeline developed by the team at Google behind Tensorflow is great achievement, as ML model design as a broad field and needs to be flexible; catering for this flexibility is no easy task. The framework needs to interpret and reimplement a model consisting of many layers and potentially millions of parameters in an entirely new context for it work in the browser. This flexibility comes at a performance cost however which can be observed in a naive implementation of \textit{Learning to See} which will be covered in more detail later. In order to recreate a model based on the provided \verb|model.json| file, separate programs in the chosen backend (WebGL, WebAssembly, WebGPU, etc.) are created to take care of specific tasks such as 2D convolutions, batchnorm, max pooling etc. Moving data through the computational graph is now the difficult part as each of these stages exists as a discrete routine which performs efficient calculations, but getting the data out of the GPU to pass on to the next program comes at a great performance cost and increases latency of the overall system, particularly with web technologies\cite{optimizing_webgl}.

@@

It was found that for a model that takes an input image size of 256$\times$256 pixels and produces an output of the same size, a dataset of 800--1000 images uniformly shuffled with a batch size of 4, trained for 200--250 epochs was sufficient to produce \textit{interesting} results --- in this case \textit{interesting} means: varied enough to provide unique experiences whilst producing a output which is representative of the input data. The models were trained on a NVIDIA Quadro RTX 6000 with 24GB memory locally, or a Tesla P100-PCIE with 16GB memory using Google Colab. In either case a usable model will train in a few hours allowing for a relatively quick development cycle. More details on the training process and some tools developed to make the process easy and fast in appendix \ref{appendix:training}.

The Pix2Pix generator is an Autoencoder similar to the ``U-Net'' architecture\cite{1505.04597} but differs in some key ways, and we offer some further changes to drastically reduce the size. The generator consists of 8 2D convolution layers downsampling the input data in width and height, and 7 transpose convolutions upsampling the data back to original size. Each layer is batchnormalized\cite{1502.03167} apart from the first, and layers 8--10 have dropout\cite{JMLR:v15:srivastava14a} applied at a rate of 50\%. The Pix2Pix architecture includes no avg- or max-pooling as all filter kernels are 4$\times$4 and a stride of 2 is used. This has the effect of downsampling the layer input by half in width and height, thus circumventing the need for any kind of pooling, otherwise known as an \textit{all convolutional net}\cite{1412.6806}. The difference proposed in this work is to greatly reduce the number of filters used in each layer. The majority of trainable parameters are found in the middle of the generator as a result of the structure described. By reducing the number of filters in each layer by up to a factor 8, very usable results were obtained whilst drastically reducing the model size.

@@

Therefore the highest compression of \verb|uint8_affine_quantize| can be used to reduce the size by an order of magnitude, while maintaining only a very slightly compromised output.

Finally, a bespoke WebGL wrapper consisting of a number of shader programs was created to handle the input data (pre-augmentation to match that in training) and the output data to render the image to the screen as video. It was found that the image handling capabilities of TensorflowJS seem to be tailored to static images. In inference, the data is of the \textit{tensor} data structure and is transformed and manipulated by the model, but at each iteration of the draw cycle a frame of video must be preprocessed and then converted into a format which can be digested by the model, and then the output of the model must be converted back into a data structure which the browser can handle to display as an image, or in our case, video. The internal computation at inference is handled by WebGL shader programs compiled by TensorflowJS, but the input/output was a rather slow process of converting the tensor data back to pixels value-by-value. The input shader programs take data from the webcam as a 16$\times$16 image and scales it up to 256$\times$256 while applying three gaussian blurs --- slightly different to the training process but the effect is the same. The data is extracted using \verb|<WebGL>.readPixels()| is then converted to a tensor object and passed to the model to process. The output of the model is a tensor object but already of the structure of a 3-channel image as a result of the architecture of the model itself. The data is easily extracted as a 32 bit float array and passed into a very simple shader program as an RGB texture to be rendered to a canvas element in the browser.

@@

\begin{small}
	\begin{center}
		\begin{tabularx}{\textwidth}{
				>{\raggedleft\arraybackslash}X
				>{\raggedleft\arraybackslash}X
				>{\raggedleft\arraybackslash}X
				>{\raggedleft\arraybackslash}X
				>{\raggedleft\arraybackslash}X
				>{\raggedleft\arraybackslash}X
			}
			\toprule
			Model  & HDF5 Size (mb) & Uint8 Compressed Size (mb) & Compression & Total Parameters & Upload Time (s) \\
			\midrule
			Small  & 25.3           & 6.3                        & 25.2\%      & 6,297,491        & 8.3             \\
			Medium & 67.3           & 16.9                       & 25.0\%      & 16,786,211       & 9.9             \\
			Large  & 217.8          & 54.5                       & 25.0\%      & 54,423,811       & 27.5            \\
			\bottomrule
		\end{tabularx}
	\end{center}
\end{small}

@@

In the table below \textit{D-GPU} means \textit{Discrete-GPU} and \textit{I-GPU} means \textit{Integrated-GPU}. \textit{FPS} is rounded to the nearest number.

\begin{small}
	\begin{center}
		\newcolumntype{b}{>{\raggedright\arraybackslash\hsize=0.65\hsize}X}
		\newcolumntype{s}{>{\raggedleft\arraybackslash\hsize=0.15\hsize}X}
		\begin{tabularx}{0.8\textwidth}{bsss}
			\toprule
			\multicolumn{4}{c}{Performance (FPS) on Various Computers} \\
			\toprule
			Machine                     & Small & Medium & Large       \\
			\midrule
			System76 Laptop 2020, D-GPU & TODO  & TODO   & TODO        \\
			System76 Laptop 2020, I-GPU & 26    & 12     & 0           \\
			Macbook Pro 2017, D-GPU     & 40    & 27     & 0           \\
			Macbook Pro 2017, I-GPU     & 26    & 14     & 0           \\
			Macbook Pro 2020, I-GPU     & 25    & 15     & 1           \\
			Macbook Pro 2015, I-GPU     & 30    & 18     & 0           \\
			MSI Laptop 2016, I-GPU      & 18    & 12     & 1           \\
			Pixel 3A Phone              & 3     & -      & -           \\
			Samsung Phone               & 4     & -      & -           \\
			\bottomrule
		\end{tabularx}
	\end{center}
\end{small}

In general the results are very promising. The average frame rate on an integrated GPU is $\sim$14FPS which suggests that the average `everyday' computer will be able to run the \textit{medium} sized model at a useable speed. However a good result in our case in more dependent on the quality of the interaction.

@@

\section{Discussion}
% As long as you can cope with
The adoption of ML research into the art world offers a reframing of the application of ML research, which in turn reframes the approach to research itself. The U-Net style, all convolutional network used for the generator in the Pix2Pix model is a particularly graceful NN which at its core is \textit{only} a series 2D convolutions (omitting batchnormalisation and skip connections). This work shows that the network itself need not be complex or large to still produce results which, to the scientific community may not be progressive or impressive, but to get this kind of work in any way to more people opens up many possibilities. As stated, if this work reaches and excites people who previously saw generative ML as off limits, then it is considered a success. Beyond that this work could lead to the development even more artistic tools in the same way that ML5\cite{ml5}, RunwayML\cite{runway} and Wekinator\cite{wekinator} have become tools for experimentation among artists and creators.

Given the smaller scale model with this workflow, the development process becomes faster and less daunting, allowing one to engage with the entire process more fully. Neural networks as a \textit{tool}, just like a paintbrush or a camera, is an exciting future in art. The iterative process involved in this work meant creating many datasets and training many models and assessing the outcome qualitatively, this helps develop a sense of \textit{how to model will react to a certain dataset}, and equally an \textit{intuition of the output}. The ability to iterate quickly is how, as a practitioner or an artist, one develops an intuition of ones tools and thus how one learns to control the medium --- a painter is not born with the ability to accurately mix color and control a brush, but it is learnt and honed over time.

This work is not only about creating art, or tools for artists. Democratising the process from the start-to-finish is important to making the process accessible to all. We currently live with machine learning and artificial intelligence; as there appears to be no limits to the possibilities there is no doubt that it will find its way into even more aspects of our lives. Therefore it is imperative that the field of machine learning becomes more accessible and transparent. One does not need to know how an engine works in order to drive a car; just like the first automobile was met with distrust[\emph{*find citation}] seen as magic, the hope of this work is to dispel those myths by giving the user their first ride in the car.

@@

We are also yet to see the perfect means of NN deployment on the internet. TensorflowJS provides great flexibility, but at some performance cost. Tools are being developed to allow internet browsers the ability to utilise low level performance enhancements specific to ML such as XNNPACK\cite{XNNPACK} while SIMD operations are now available to WebAssembly\cite{SIMD} --- both features being utilised to perform background blurring and substitution in-browser and on mobile\cite{google_background}. WebGPU\cite{webgpu} and WebNN\cite{webnn} also lie on the horizon showing great promise as web specific ML tools able to handle the data structures of ML. As it stands however WebGL remains the strongest web API to interface with the GPU and with simple models such as the \textit{all convolutional net} it is simply a matter of rethinking tensor data structure into 2D textures to allow for efficient GLSL shader programs to operate on the data. A sample NN of similar structure to the Pix2Pix generator was created to test this hypothesis and shows potential. Early tests show that the running time of a full size Pix2Pix generator model on an integrated GPU is the same as the TensorflowJS model with the WebGL wrapper as used in this work. However keeping the data structures suited to WebGL means it is trivial to display the weights and biases and possibly create novel interactions with even deeper aspects of the model --- more details can be seen in appendix \ref{appendix:webgl_cnn}.

@@

\begin{appendices}
	\section{Training Process}
	\label{appendix:training}
	Some text that will be the appendix.

@@

\end{appendices}

>>>>

\begin{abstract}
	% 250 Words
	The lines between scientific researcher and artist are continually being blurred as artists adopt machine learning methods\cite{anna_ridler}, or researchers apply their tools in novel and interesting ways\cite{deep_dream_guy} rethinking how cutting-edge research can be applied to create image generation tools or interrogate the tools themselves. \textit{Learning to See} by Memo Akten\cite{lts_memo} is one such seminal artwork which does both. Reframing the tools which we are in fact living with day-to-day in the context of the gallery provides an opportunity to think deeper about how a neural network interprets the data it is given and to witness the capabilities and limitations of such a system. \textit{Learning to See} in particular has a uniquely simple interaction allowing the visitor to play with the network in real time. This work shows that simply shrinking the Pix2Pix\cite{1611.07004} model which powers \textit{Learning to See}, and a custom made WebGL wrapper for TensorflowJS to make a pipeline more suited to video achieves useable results when served online, even on commonplace integrated graphics processing units. This brings the unique interaction developed by Akten \textit{et al}\cite{2003.00902} into the homes of many people and offers exciting possibilites to future video based interaction with machine learing on the internet. The work is currently online here: \url{https://learning-to-learn-to-see.netlify.app/}, and code available here: \url{https://github.com/joshmurr/cci-pix2pix-demo/}.
\end{abstract}

@@

\section{Introduction}
% 500 Words
% --- 568 WORDS ---
Machine Learning (ML) is a part of our everyday lives. For the technical among us it is currently easy enough to see the patterns in the outputs of these black-box systems---as the end of our sentence is predicted as we type, or an item we did not know we wanted is offered to us as we shop. There are however many non-technical users of ML systems who can only see the black-box, or perhaps nothing at all; as Arthur C. Clarke famously said \textit{``Any sufficiently advanced technology is indistinguishable from magic.''}\cite{toffler_1972} It is often the case that exciting new technologies are adopted by artists as readily as industry because the gallery provides a unique environment to present back the to the user one of these black-boxes in a drastically different context. Artists such as Mario Klingemann\cite{mario_klingemann}, Terence Broad\cite{terry_broad} or Anna Ridler\cite{anna_ridler} are all excellent examples of artists who have done just this.

\textit{Learning to See} by Memo Akten\cite{lts_memo} is another example, but unique in its own way in that it gives the user an interactive experience of the artwork; an immediate image translation of arbitrary objects placed on a tabletop to a rich image of waves crashing, the night sky or ever changing flowers. The user is able to manipulate the input data and witness how the model interprets this to produce an output image in front of them. Although the model still exists as a black box, being able to see how input-affects-output in real time is extremely powerful.

@@

Unfortunately the gallery in itself is a black box to many, entry restricted by a fee or simply by the location of the gallery. Moving such interactive experiences online is a clear next step as once something is online it is immediately available to anyone with an internet connection --- this then poses numerous challenges. Even a modestly sized NN can contain many millions of parameters. Modern computers are able to perform millions of floating point operations per second (FLOPS) but advances in ML research have shown that \textit{deeper} models and more data produces better results\cite{2005.14165}; as such, without expensive hardware acceleration from discrete graphics processing units (GPU) or even more modern tensor processing units (TPU), an average modern computer cannot crunch the numbers needed to run a large, deep NN\@. Even if one \textit{could} perform the necessary FLOPS, storing this number of parameters can take up gigabytes of memory and so downloading the weights and biases of a pre-trained model can be a time consuming process detracting from the real-time focus.

@@

\textit{Learning to See} is an adaptation of the Pix2Pix model\cite{1611.07004} which is a Conditional Generative Adversarial Network (cGAN)\cite{1701.00160}. Unlike a conventional GAN\cite{1406.2661}, a cGAN takes in two data samples as input, in the case of the Pix2Pix model it is a \textit{target} image and an \textit{input} image. Pix2Pix also learns a structured loss which means it is able to penalize any possible structure that differs between output and target\cite{1611.07004}. This model architecture is extremely flexible and has been shown to be useful for many use cases as detailed in the original paper. However neither the original Pix2Pix paper nor \textit{Learning to See} had an incentive to create a small model; a good result is simply a sharp image with a predictable output from a given input. As stated by the Pix2Pix authors measuring the results of an image generating GAN is an open issue\cite{1606.03498, 1611.07004} and in an artistic context quantifiying the outcome is largely pointless --- a good result is one which delivers the message of the artist.

@@

Designing and training a NN is of course not limited to the big name ML frameworks like Tensorflow\cite{tf} or PyTorch\cite{pytorch}, but given the modular nature of a NN and the want in most cases to utilise the GPU for faster training and inference, it is no surprise that a limited number of ML frameworks have become the go-to to create and iterate on a novel ML model and to take care of interfacing with a highly optimized backend and the GPU at a lower level. Tensorflow offers a relatively full end-to-end pipeline from designing and training a model in Python, to saving and converting a model into something can be run in the browser using TensorflowJS\cite{tfjs}. A trained model is simply a collection of weights and biases and to use this model on a different device or context, these weights and biases need to be loaded into a model of the same architecture as was used to train the model. The pipeline from training-to-browser is largely about keeping data structures in order and consistent such that the next process knows how to work with it. As part of the process Tensorflow has developed the TensorflowJS Converter\cite{tfjs_converter} which converts a saved model into a a series of binary files which hold the weights and a \verb|.json| file which is the blueprints of the model. When uploading a model TensorflowJS dynamically creates a succession of routines in the chosen backend (WebGL, WebAssembly, WebGPU, etc.) to take care of specific tasks such as 2D convolutions, batchnorm, max pooling etc.%  Moving data through the computational graph is now the difficult part as each of these stages exists as a discrete routine which performs efficient calculations, but getting the data out of the GPU to pass on to the next program comes at a great performance cost and increases latency of the overall system, particularly with web technologies\cite{optimizing_webgl}.

@@

It was found that for a model that takes an input image size of 256$\times$256 pixels and produces an output of the same size, a dataset of 800--1000 images uniformly shuffled with a batch size of 4 was sufficient to produce \textit{interesting} results --- in this case \textit{interesting} means: varied enough to provide unique experiences whilst producing a output which is representative of the input data. Differences between the model sizes can be seen in \ref{table:model_details}. The models were trained on a NVIDIA Quadro RTX 6000 with 24GB memory locally, or a Tesla P100-PCIE with 16GB memory using Google Colab. In either case a usable model will train in a few hours allowing for a relatively quick development cycle. %More details on the training process and some tools developed to make the process easy and fast in appendix \ref{appendix:training}.

The Pix2Pix generator is an Autoencoder similar to the ``U-Net'' architecture\cite{1505.04597} but differs in some key ways, and we offer some further changes to drastically reduce the size. The generator consists of 8 2D convolution layers downsampling the input data in width and height, and 7 transpose convolutions upsampling the data back to original size. Each layer is batchnormalized\cite{1502.03167} apart from the first, and layers 8--10 have dropout\cite{JMLR:v15:srivastava14a} applied at a rate of 50\%. The Pix2Pix architecture includes no avg- or max-pooling as all filter kernels are 4$\times$4 and a stride of 2 is used. This has the effect of downsampling the layer input by half in width and height, thus circumventing the need for any kind of pooling, otherwise known as an \textit{all convolutional net}\cite{1412.6806}. The difference proposed in this work is to greatly reduce the number of filters used in each layer. The majority of trainable parameters are found in the middle of the generator as a result of the structure described. By reducing the number of filters in each layer by up to a factor of 8, very usable results were obtained whilst drastically reducing the model size --- more discussion on the impacts of a smaller model are found in Results.

@@

Therefore the highest compression of \verb|uint8_affine_quantize| can be used to reduce the size by 4$\times$, while maintaining only a very slightly compromised output.

Finally, a bespoke WebGL wrapper consisting of a number of shader programs was created to handle the input data (pre-augmentation to match that in training) and the output data to render the image to the screen as video. It was found that the image handling capabilities of TensorflowJS seem to be tailored to static images. In inference, the data is of the \textit{tensor} data structure and is transformed and manipulated by the model, but at each iteration of the draw cycle a frame of video must be preprocessed and then converted into a format which can be digested by the model, and then the output of the model must be converted back into a data structure which the browser can handle to display as an image, or in our case, video. The internal computation at inference is handled by WebGL shader programs compiled by TensorflowJS, but the input/output was a rather slow process of converting the tensor data back to pixels value-by-value. The input shader programs receives data from the webcam as a 16$\times$16 image and scales it up to 256$\times$256 while applying three gaussian blurs --- slightly different to the training process but the effect is the same. The data is extracted using \verb|<WebGL>.readPixels()|, converted to a tensor object and passed to the model to process. The output of the model is a tensor object but already of the structure of a 3-channel image as a result of the architecture of the model itself. The data is easily extracted as a 32 bit float array and passed into a very simple shader program as an RGB texture to be rendered to a canvas element in the browser.

@@

\begin{small}
	\begin{center}
		\newcolumntype{b}{>{\raggedleft\arraybackslash}X}
		\begin{tabularx}{\textwidth}{bbbbbbb}
			\toprule
			Model  & HDF5 Size (mb) & Uint8 Compressed Size (mb) & Compression & Total Parameters & Epochs Trained & Upload Time (s) \\
			\midrule
			Small  & 25.3           & 6.3                        & 25.2\%      & 6,297,491        & 80             & 8.3             \\
			Medium & 67.3           & 16.9                       & 25.0\%      & 16,786,211       & 120            & 9.9             \\
			Large  & 217.8          & 54.5                       & 25.0\%      & 54,423,811       & 200            & 27.5            \\
			\bottomrule
		\end{tabularx}
		\label{table:model_details}
	\end{center}
\end{small}

@@

In the table below \textit{D-GPU} means \textit{Discrete-GPU} and \textit{I-GPU} means \textit{Integrated-GPU}. \textit{FPS} is rounded to the nearest number. All figures are taken from the online platform.

\begin{small}
	\begin{center}
		\newcolumntype{b}{>{\raggedright\arraybackslash\hsize=0.65\hsize}X}
		\newcolumntype{s}{>{\raggedleft\arraybackslash\hsize=0.15\hsize}X}
		\begin{tabularx}{0.8\textwidth}{bsss}
			\toprule
			\multicolumn{4}{c}{Performance (FPS) on Various Computers} \\
			\toprule
			Machine                     & Small & Medium & Large       \\
			\midrule
			System76 Laptop 2020, D-GPU & 50    & 41     & 10          \\
			System76 Laptop 2020, I-GPU & 26    & 12     & 0           \\
			Macbook Pro 2020, I-GPU     & 25    & 15     & 1           \\
			Macbook Pro 2017, D-GPU     & 40    & 27     & 0           \\
			Macbook Pro 2017, I-GPU     & 26    & 14     & 0           \\
			Macbook Pro 2015, I-GPU     & 30    & 18     & 0           \\
			Macbook Pro 2015, D-GPU     & 27    & 18     & 1           \\
			MSI Laptop 2016, I-GPU      & 18    & 12     & 1           \\
			Pixel 3A Phone              & 3     & -      & -           \\
			Samsung Galaxy A20e Phone   & 4     & -      & -           \\
			\bottomrule
		\end{tabularx}
	\end{center}
\end{small}

In general the results are very promising. The average frame rate on an integrated GPU is $\sim$14FPS which suggests that the average `everyday' computer will be able to run the \textit{medium} sized model at a useable speed while a gaming-laptop with a discrete GPU can observe good results even on the largest model.

@@

\section{Discussion}
% As long as you can cope with
The adoption of ML research into the art world offers a reframing of the application of ML research, which in turn reframes the approach to research itself. The U-Net style, all convolutional network used for the generator in the Pix2Pix model is a particularly graceful NN which at its core is \textit{only} a series 2D convolutions (omitting batchnormalisation and skip connections). This work shows that the network itself need not be complex or large to still produce results which are useable, visually interesting and accessible. As stated, if this work reaches and excites people who previously saw generative ML as off limits, then it is considered a success. Beyond that this work could lead to the development even more artistic tools in the same way that ML5\cite{ml5}, RunwayML\cite{runway} and Wekinator\cite{wekinator} have become tools for experimentation among artists and creators.

Given the smaller scale model with this workflow, the development process becomes faster and less daunting, allowing one to engage with the entire process more fully. Neural networks as a \textit{tool}, just like a paintbrush or a camera, is an exciting future in art. The iterative process involved in this work meant creating many datasets and training many models and assessing the outcome qualitatively, this helps develop a sense of \textit{how the model will react to a certain dataset}, and equally an \textit{intuition of the output}. The ability to iterate quickly is how, as a practitioner or an artist, one develops an intuition of ones tools and thus how one learns to control the medium --- a painter is not born with the ability to accurately mix color and control a brush, but it is learnt and honed over time.

This work is not only about creating art, or tools for artists. Democratising the process from start-to-finish is important to making the process accessible to all. We currently live with machine learning and artificial intelligence; as there appears to be no limits to the possibilities there is no doubt that it will find its way into even more aspects of our lives. Therefore it is imperative that the field of machine learning becomes more accessible and transparent. The first automobile was met with distrust[\emph{*find citation}] and seen as magic, the hope of this work is to dispel some myths of ML by giving the user their first ride in the car.

@@

We are also yet to see the perfect means of NN deployment on the internet. TensorflowJS provides great flexibility, but at some performance cost. Tools are being developed to allow internet browsers the ability to utilise low level performance enhancements specific to ML such as XNNPACK\cite{XNNPACK} while SIMD operations are now available to WebAssembly\cite{SIMD} --- both features being utilised to perform background blurring and substitution in-browser and on mobile\cite{google_background}. WebGPU\cite{webgpu} and WebNN\cite{webnn} also lie on the horizon showing great promise as web specific ML tools able to handle the data structures of ML. As it stands however WebGL remains the strongest web API to interface with the GPU and with simple models such as the \textit{all convolutional net} it is simply a matter of rethinking tensor data structure into 2D textures to allow for efficient GLSL shader programs to operate on the data. A sample NN of similar structure to the Pix2Pix generator was created to test this hypothesis and shows potential. Early tests show that the running time of a full size Pix2Pix generator model on an integrated GPU is the same as the TensorflowJS model with the WebGL wrapper as used in this work. The advantages of keeping the data structures suited to WebGL means it is trivial to display the weights and biases and possibly create novel real-time interactions with even deeper aspects of the model.

@@

\begin{comment}
\begin{appendices}
	\section{Training Process}
	\label{appendix:training}
	Some text that will be the appendix.

@@

\end{appendices}
\end{comment}
